Estimating the association between Facebook adoption and well-being in 72 countries

### 2022-10-20 ms.Rmd
Submitted

---
title: Estimating the association between Facebook adoption and well-being in 72 countries

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
author:
  - name: Matti Vuorre
    affiliation: a
  - name: Andrew K. Przybylski
    affiliation: a,1
address:
  - code: a
    address: Oxford Internet Institute, University of Oxford, 1 St Giles, Oxford OX1 3JS, United Kingdom

corresponding_author:
  code: 1
  text: "To whom correspondence should be addressed. E-mail: andy.przybylski@oii.ox.ac.uk"

# For footer text
lead_author_surname: Vuorre

author_contributions: |
  The authors made the following contributions. Matti Vuorre: Design, Administration, Analysis, Visualization, Writing; Andrew K. Przybylski: Design, Administration, Funding, Resources, Super- vision, Visualization, Writing.

## Remove this if not required
conflict_of_interest: |
  The authors declare no conflicts of interest.

abstract: |
  Social media's potential effects on well-being have received considerable research interest, but much of past work is hampered by an exclusive focus on demographics in the Global North and inaccurate self-reports of social media engagement. We describe associations linking 72 countries' Facebook adoption to the well-being of 946,798 individuals from 2008 to 2019. We found no associations suggesting that the global penetration of social media has caused widespread psychological harm: Facebook adoption predicted life satisfaction and positive experiences positively, and negative experiences negatively, both between countries and within countries but over time, and the associations were small. Facebook adoption was more positively associated with well-being for younger individuals, but country-specific results were mixed. In order to move beyond studying aggregates and to better understand social media's causal effects, we need more transparent collaborative research between independent scientists and the technology industry.

significance: |
  Billions of individuals across the globe use social media on a regular basis. The widespread adoption of social media---most prominently Facebook---has fundamentally shifted how humans conduct their lives. Consequently, researchers, healthcare professionals, and policy makers have repeatedly raised alarms that the spread of social media may be negatively affecting individuals', particularly young ones', mental health and well-being on a global scale. We examined the plausibility of this idea by associating 72 nations' yearly Facebook adoption rates to the well-being of 946,798 individuals from 2008 to 2019. Overall, we found that increased rates of daily and monthly active Facebook users slightly predicted greater levels of well-being, when comparing countries to other countries, or countries to themselves but over time. These findings should guide debate surrounding social media towards more empirical foundations: The importance of this topic requires that independent scientists gain access to more detailed data from the technology industry in order to better determine how, when, and why modern online platforms might be affecting their users.

# acknowledgements: |
#   Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.

keywords:
  - well-being
  - social media

## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle

bibliography: references.bib
csl: pnas.csl

## change to true to add optional line numbering
lineno: true

output: rticles::pnas_article
---

\raggedbottom

@@

# Plotting options
theme_set(
  theme_linedraw(
    base_size = 8
  ) +
    theme(
      panel.grid = element_blank(),
      strip.text = element_text(margin = margin(4, 4, 4, 4, "pt")),
      axis.text = element_text(size = rel(0.7)),
      axis.ticks = element_line(size = rel(0.4))
    )
)
```

@@

```{r coefficients-calculate}
path <- "models/coefficients.rds"
if (!file.exists(path)) {
  # Calculate contrasts for each outcome
  h_dau <- c(
    "Average" = "dau_cw = 0",
    "13-34_Female" = 
      "dau_cw + dau_cw:age1 * -1 + dau_cw:sex1 * -1 + dau_cw:age1:sex1 * 1 = 0",
    "13-34_Male" = 
      "dau_cw + dau_cw:age1 * -1 + dau_cw:sex1 * 1 + dau_cw:age1:sex1 * -1 = 0",
    "35+_Female" = 
      "dau_cw + dau_cw:age1 * 1 + dau_cw:sex1 * -1 + dau_cw:age1:sex1 * -1 = 0",
    "35+_Male" = 
      "dau_cw + dau_cw:age1 * 1 + dau_cw:sex1 * 1 + dau_cw:age1:sex1 * 1 = 0",
    "13-34" = "dau_cw + dau_cw:age1 * -1 = 0",
    "35+" = "dau_cw + dau_cw:age1 * 1 = 0",
    "Female" = "dau_cw + dau_cw:sex1 * -1 = 0",
    "Male" = "dau_cw + dau_cw:sex1 * 1 = 0",
    "Between-country" = "dau_cb = 0",
    "X by age" = "dau_cw:age1 = 0",
    "X by sex" = "dau_cw:sex1 = 0",
    "X by age by sex" = "dau_cw:age1:sex1 = 0"
  )
  h_mau <- c(
    "Average" = "mau_cw = 0",
    "13-34_Female" = 
      "mau_cw + mau_cw:age1 * -1 + mau_cw:sex1 * -1 + mau_cw:age1:sex1 * 1 = 0",
    "13-34_Male" = 
      "mau_cw + mau_cw:age1 * -1 + mau_cw:sex1 * 1 + mau_cw:age1:sex1 * -1 = 0",
    "35+_Female" = 
      "mau_cw + mau_cw:age1 * 1 + mau_cw:sex1 * -1 + mau_cw:age1:sex1 * -1 = 0",
    "35+_Male" = 
      "mau_cw + mau_cw:age1 * 1 + mau_cw:sex1 * 1 + mau_cw:age1:sex1 * 1 = 0",
    "13-34" = "mau_cw + mau_cw:age1 * -1 = 0",
    "35+" = "mau_cw + mau_cw:age1 * 1 = 0",
    "Female" = "mau_cw + mau_cw:sex1 * -1 = 0",
    "Male" = "mau_cw + mau_cw:sex1 * 1 = 0",
    "Between-country" = "mau_cb = 0",
    "X by age" = "mau_cw:age1 = 0",
    "X by sex" = "mau_cw:sex1 = 0",
    "X by age by sex" = "mau_cw:age1:sex1 = 0"
  )
  coefs <- fits %>% 
    select(-data, -model) %>% 
    mutate(h = if_else(predictor == "dau", list(h_dau), list(h_mau))) %>% 
    mutate(
      world = map2(
        fit, h,
        ~hypothesis(.x, .y) %>% 
          .[["hypothesis"]] %>%
          tibble()
      ),
      country = map2(
        fit, h,
        ~hypothesis(.x, .y, scope = "coef", group = "country") %>%
          .[["hypothesis"]] %>%
          tibble()
      )
    ) %>% 
    select(-fit, -h)

  # Both predictor and outcome are 0-1, so scale can be interpreted as % change as a function of % change

  coefs <- coefs %>% 
    pivot_longer(c(world, country)) %>% 
    unnest(value) %>% 
    mutate(
      country = if_else(name == "country", as.character(Group), "World")
    ) %>% 
    # Between country contrast not country-specific!
    filter(!(Hypothesis == "Between-country" & country != "World")) %>% 
    # Neatly formatted result string
    mutate(
      across(c(Estimate, CI.Lower, CI.Upper), list("str" = ~number(., .01))),
      result = str_glue("{Estimate_str}% [{CI.Lower_str}, {CI.Upper_str}]")
    )

@@

To better understand the plausible range of associations, we linked data tracking Facebook's global adoption with three indicators of well-being. We examined 72 countries' per capita daily active Facebook users (DAU) in two age brackets (13-34 and 35+ years) as predictors of life satisfaction (LS), negative (NE), and positive psychological experiences (PE) at the level of years spanning 2008 to 2019. The well-being data represented 946,798 individuals' responses from the nationally representative Gallup World Poll Survey [@gallupHowDoesGallup2014]. We joined these unique datasets to conduct a descriptive study to answer three basic yet important questions: First, to what extent is Facebook adoption associated with well-being? Second, do these associations differ by age or sex. And finally, how might these associations have differed between countries?

# Average within- and between-country associations
tab_world <- coefs %>% 
  filter(
    name == "world",
    Hypothesis %in% c("Average", "Between-country")
  ) %>% 
  mutate(
    outcome = case_when(
      outcome == "Life_satisfaction" ~ "LS",
      outcome == "Negative_experiences" ~ "NE",
      outcome == "Positive_experiences" ~ "PE"
    )
  ) %>% 
  select(
    predictor, Outcome = outcome, Hypothesis, result
  ) %>% 
  pivot_wider(names_from = Hypothesis, values_from = result) %>% 
  rename(
    `Within country` = Average,
    `Between country` = `Between-country`
  ) %>% 
  select(predictor, Outcome, `Between country`, `Within country`) %>% 
  arrange(predictor, Outcome)

tab_world %>% 
  filter(predictor == "dau") %>% 
  select(-predictor) %>% 
  mutate(
    across(
      ends_with("country"), 
      ~if_else(str_starts(., "-"), ., paste0("\\phantom{-}", .))
    )
  ) %>% 
  apa_table(
    caption = "Average Facebook DAU and well-being associations",
    font_size = "small",
    note = "LS: Life satisfaction, NE: Negative experiences: PE: Positive experiences. Numbers indicate posterior means and 95\\%CIs.",
    span_text_columns = FALSE,
    escape = FALSE,
    placement = "htb",
    align = 'llll'
  )
```

# Percentages of countries where DAU was over 25% by year
tmp <- d_fb %>% 
  group_by(year, age) %>% 
  summarise(median_dau = percent(median(dau))) %>% 
  pivot_wider(names_from = age, values_from = median_dau)
tmp
```

Facebook adoption increased markedly from 2008 when the median per capita DAU across these 72 countries was `r tmp[1,2]` (ages 13-34) and `r tmp[1,3]`% (35+), to `r tmp[12,2]` (13-34) and `r tmp[12,3]` (35+; Figure 1A) in 2019. During this period Facebook adoption by younger individuals plateaued before saturation in many countries, but not for older individuals. At the same time, we did not observe correspondingly large and uniform changes across measures of well-being (Figure 1B) [@vuorreGlobalWellbeingMental2022].

To answer our first question of how Facebook adoption relates to well-being, we specified a multilevel linear meta-regression model of each well-being outcome on time, sex, age, and DAU (both within- and between-country-centered, see SI). We included all interactions between sex, age, and time, and sex, age, and DAU (within-country-centered), and allowed all parameters to vary randomly across countries. We first examined whether the relative standing of countries on their average Facebook adoption predicted well-being (the model's between-country associations). We found that countries with greater average DAU had somewhat higher levels of life satisfaction and positive experiences, and lower levels of negative experiences than countries with lower DAU (Table 1). However, there are large and important differences between countries in factors that might underlie differences in both social media adoption and well-being, such as socioeconomic conditions. While descriptively informative, these associations are likely to indicate between-country confounding factors.

```{r fig-data, fig.pos = "H"}
#| include: true
#| fig.height: 3.6
#| fig.width: 3.5
#| fig.cap: "A. 72 countries' daily active Facebook users in two age brackets from 2008 to 2019. (Percentages may exceed 100\\% due to inaccurate estimation of either DAU or population size.) B. The same 72 countries' mean responses to three well-being scales in the Gallup World Poll from 2008 to 2019, separated by age category and sex. LS: Life satisfaction, NE: Negative experiences: PE: Positive experiences."

# Figure of all variables over time
p_a <- d_fb %>% 
  ggplot(aes(year, dau, group = country)) +
  scale_y_continuous(
    "Percentage of population", 
    expand = expansion(.03),
    breaks = extended_breaks(),
    labels = ~percent(.)
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2), 
    labels = ~str_replace(., "20", "'"),
    expand = expansion(0.01)
  ) +
  geom_line(size = .125) +
  # geom_smooth(group = 1, method = "gam", se = F) +
  facet_wrap("age", nrow = 1) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  theme(
    strip.text.x = element_text(margin = margin(2,0,2,0, "pt")),
    text = element_text(size = 5)
    )

p_b <- d %>% 
  # Exclude rows where FB data existed but not GWP data
  drop_na(val) %>% 
  # Actual lowest age in GWP
  mutate(age = factor(age, labels = c("15-34", "35+"))) %>% 
  mutate(
    outcome = case_when(
      outcome == "Life_satisfaction" ~ "LS",
      outcome == "Negative_experiences" ~ "NE",
      outcome == "Positive_experiences" ~ "PE"
    )
  ) %>% 
  mutate(panel = str_glue("{age} ({sex})")) %>% 
  ggplot(aes(year, val, group = country)) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(
    "Value", 
    expand = expansion(.01),
    breaks = extended_breaks(),
    labels = ~percent(.)
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2),
    labels = ~str_replace(., "20", "'")
  ) +
  geom_line(size = .1) +
  facet_grid(
    outcome~panel, 
    scales = "free_y",
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  theme(
    strip.text.x = element_text(margin = margin(2,0,2,0, "pt")),
    strip.text.y = element_text(margin = margin(0,2,0,2, "pt")),
    legend.position = "none",
    text = element_text(size = 5)
  )

(p_a / p_b) +
  plot_layout(heights = c(3, 7)) +
  plot_annotation(tag_levels = "A")
```

Variations in such confounding factors are likely to be significantly smaller within countries over time. We therefore next focused on the model's within-country associations, which measure the extent to which Facebook adoption in a given country predicted well-being in that country adjusting for general linear temporal trends in the well-being outcome. On average across age, and sex, we found that years with greater than the country's average DAU were associated with greater levels of life satisfaction and positive experiences, and lower levels of negative experiences (Table 1), although the first association's 95%CI contained zero. For the average country in our sample, Facebook adoption positively predicted well-being. The magnitudes of these associations were small; a one-percent increase in DAU predicted a `r tab_world[3,4]` increase in positive experiences for the average country. While these associations indicate relations within countries and adjust for confounders that vary linearly with time by including time as a predictor, they are still susceptible to confounders and do not indicate causal relations. Rather, this association describes that, all else being equal, years with greater Facebook adoption tended to be those with greater levels of well-being.

```{r demographics}
coefs %>% 
  filter(
    str_detect(Hypothesis, " by "), 
    name == "world"
  ) %>% 
  select(
    Predictor = predictor, Outcome = outcome, Hypothesis, result
  ) %>% 
  pivot_wider(names_from = Outcome, values_from = result)
coefs %>% 
  filter(
    Hypothesis %in% c("13-34", "35+", "Female", "Male"),
    name == "world"
  ) %>% 
  select(
    Predictor = predictor, Outcome = outcome, Hypothesis, result
  ) %>% 
  pivot_wider(names_from = Outcome, values_from = result)

tab_demo <- coefs %>% 
  filter(
    str_detect(Hypothesis, "_"),
    name == "world"
  ) %>% 
  select(
    Predictor = predictor, Outcome = outcome, Hypothesis, result
  ) %>% 
  pivot_wider(names_from = Outcome, values_from = result)
tab_demo
```

We then turned to our second question and assessed whether within-country associations linking Facebook adoption to well-being differed between age and sex. Differences in the associations were greater between the younger and older demographics than they were between females and males (Figure 2, Average). DAU predicted life satisfaction more positively in the younger age bracket, but these estimates' CIs did not exclude zero. DAU predicted negative experiences negatively only in the younger age group. These differences were less pronounced for positive experiences. These results show that Facebook adoption was, for the average nation in our sample, more positively predictive of well-being for individuals below 35, and that differences between males and females were generally smaller than those between age groups. Importantly, they did not indicate that the well-being of younger individuals is particularly linked to social media adoption.

```{r}
# Can we show regions in the forest plot as secondary axis? 
# Table of percentages of countries with credible associations
tmp <- coefs %>% 
  filter(name == "country", Hypothesis == "Average") %>% 
  group_by(predictor, outcome, Star, direction=sign(Estimate)) %>% 
  count() %>% 
  group_by(predictor, outcome) %>% 
  # mutate(n = percent(n / sum(n), 1)) %>% 
  mutate(x = interaction(Star, direction)) %>% 
  select(-Star, -direction) %>% 
  pivot_wider(names_from=x, values_from=n) %>% 
  replace_na(list(`.-1` = 0, `.1` = 0, `*.1` = 0, `*.-1` = 0))
tmp
coefs %>% 
  filter(predictor == "dau", name == "country", Hypothesis == "Average") %>% 
  filter(Star=="*") %>% 
  arrange(outcome, country) %>% 
  select(outcome, country, Estimate:CI.Upper)
```

However, while informative aggregates, these results do not describe associations between Facebook adoption and well-being for any individual country, but rather their averages. To answer our third question, we computed country-specific estimates for each sex and age group (Figure 2). For life satisfaction, `r tmp[1, 5]` countries had credibly positive average associations. For negative experiences, `r tmp[2, 6]` countries had credibly negative average association, and `r tmp[3, 5]` countries had a credibly positive average association between DAU and positive experiences. There were no credible (at the 95% level) country-level negative associations between DAU and life-satisfaction. Although many country-level estimates were uncertain, these results aligned with the general pattern of positive associations between Facebook adoption and country-level well-being.

We found that, overall, Facebook adoption positively predicted well-being. This association held when comparing countries to other countries and to themselves over time. While these descriptive results do not speak to causal effects, they align with other recent findings suggesting that technology has not become increasingly harmful over time [@vuorreThereNoEvidence2021], and that the increased adoption of Internet technologies is not, overall, associated with widespread psychological harms [@vuorreGlobalWellbeingMental2022]. That said, these conclusions are qualified because our focus on daily active users could miss those who use Facebook less regularly. To test this possibility, we also conducted our analyses using monthly active Facebook users (MAU) as the predictor. Those results were substantively the same (see online analysis supplement; <https://doi.org/10.5281/zenodo.7086277>).

```{r fig-forest, fig.pos="H"}
#| echo: false
#| include: true
#| fig.height: 4.2
#| fig.width: 3.5
#| fig.cap: "Forest plot of estimated associations between country-year level per-capita daily active Facebook users (DAU) and life satisfaction (LS), negative experiences (NE), and positive experiences (PE). Units indicate percentage change in outcome as function of percentage increase in within-country centered DAU. Filled points indicate estimates whose posterior 95\\%CI excluded zero. See the online supplementary materials for corresponding figure but with MAU as predictor."

tmp <- coefs %>%   
  filter(
    # name != "world",
    str_detect(Hypothesis, "_")
  ) %>% 
  separate(Hypothesis, c("Age", "Sex"), sep = "_") %>% 
  mutate(
    outcome = case_when(
      outcome == "Life_satisfaction" ~ "LS",
      outcome == "Negative_experiences" ~ "NE",
      outcome == "Positive_experiences" ~ "PE"
    )
  ) %>% 
  # mutate(outcome = str_glue("{outcome} ({Sex})")) %>% 
  filter(predictor == "dau")

tmp %>%
  ggplot() +
  aes(
    x = Estimate, 
    y = fct_rev(country), 
    col = Age, 
    fill = interaction(Age, Star, sep = " "), 
    shape = Sex
  ) +
  scale_y_discrete(
    "Country",
    expand = expansion(.015),
    labels = ~str_replace(., "World", "Average")
  ) +
  scale_x_continuous(
    "Association with daily active Facebook users (95%CI)",
    breaks = extended_breaks(7),
    expand = expansion(.1)
  ) +
  scale_shape_manual(
    values = c(21, 22)
  ) +
  scale_color_brewer(
    "Age", palette = "Set1", aesthetics = c("color")
  ) +
  scale_fill_manual(values = c("white", "white", "#E41A1C", "#377EB8")) +
  geom_vline(xintercept = 0, lty = 2, size = .25, col = "grey50") +
  geom_linerangeh(
    aes(xmin = CI.Lower, xmax = CI.Upper),
    size = .15, position = position_dodge2v(.7), 
    show.legend = FALSE
  ) +
  geom_point(
    size = .65, stroke = .45,
    position = position_dodge2v(.7)
  ) +
  facet_wrap(
    "outcome", scales = "free_x", nrow = 1,
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(size = 1.5)),
    shape = guide_legend(override.aes = list(size = 1.5))
  ) +
  theme(
    text = element_text(size = 5),
    strip.text.x = element_text(margin = margin(2,0,2,0, "pt")),
    axis.text.y = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(
      size = .1, linetype = 2, color = "gray60"
    ),
    legend.position = "bottom", 
    legend.box.margin = margin(-15, 0, -10, 0, "pt"),
    legend.background = element_rect(fill = NA),
    legend.key = element_rect(fill = NA),
    legend.text = element_text(size = rel(0.9))
  )
```

It is widely accepted that social media and the Internet more broadly have changed how humans socialize, organize, and seek leisure, but it is not obvious or necessary that their wide adoption has influenced well-being. In this study we used the best data available to describe how global Facebook adoption relates to well-being across 72 countries over a 12-year period. We found generally positive associations between Facebook adoption and well-being which were partially qualified by demographics and not uniform across countries.

We did not find evidence that increased social media adoption is consistently negatively associated with well-being. Nevertheless, our analyses cannot and do not rule out that possibility because more fine-grained data needed to demonstrate this more conclusively either do not exist or are not available. If we are to move past description to prediction or intervention, independent scientists and online platforms will need to collaborate in new, transparent ways. As it stands now, only a handful of scientists working in the technology industry have the data required to advance this line of inquiry. If we are to understand and improve well-being in the digital age, this must change.

# Supporting information

@@

Both datasets (Facebook and Gallup) are proprietary and we therefore could not share them with this manuscript. Our analytic code, along with synthetic datasets and an online analysis supplement, is available at <https://doi.org/10.5281/zenodo.7086277>. Researchers can contact Facebook (ccobb@fb.com) to reproduce our analyses with the actual Facebook adoption dataset. The Gallup data are available online to subscribing institutions.

@@

Before analysing the data, we converted Facebook DAU and MAU to proportions of each country's and age group's yearly population sizes using population data from the United Nations Department of Economic and Social affairs (<https://population.un.org/wpp/Download/Standard/Population/>).

We examined the association between Facebook adoption and well-being through bayesian hierarchical regression models, estimated separately for DAU and MAU, and each of the three well-being outcomes. We regressed the outcome $y$ on time (decades, centered on 2014), within-country centered $x$ (DAU / MAU), age, sex, all the interactions except time by $x$, and the between-country centered DAU. We allowed all coefficients to vary randomly across the 72 countries, except the between-country predictor. Because we aggregated the outcomes, we included to standard errors of the outcome in the model to account for the varying group sizes. Formally, we specified the model as

>>>>

---
title: Estimating the association between Facebook adoption and well-being in 72 countries
shorttitle: Facebook and well-being
leftheader: Facebook and well-being
author: 
  - name: Matti Vuorre
    affiliation: 1
    address: Oxford Internet Institute, University of Oxford
    email: matti.vuorre@oii.ox.ac.uk
    role:
      - Design
      - Administration
      - Analysis
      - Visualization
      - Writing
  - name: Andrew K. Przybylski
    affiliation: 1
    corresponding: yes
    address: Oxford Internet Institute, University of Oxford
    email: andy.przybylski@oii.ox.ac.uk
    role:
      - Design
      - Administration
      - Funding
      - Resources
      - Supervision
      - Visualization
      - Writing

affiliation:
  - id: 1
    institution: Oxford Internet Institute, University of Oxford

authornote: |
  \noindent \textbf{This manuscript is not yet peer-reviewed.}

  This research was supported by the Huo Family Foundation and Economic and Social Research Council Grant ES/T008709/1.

  This study and methodology was reviewed and approved by the Oxford Internet Institute's Departmental Research Ethics Committee (SSH_OII_CIA_21_084). The authors declare no conflicts of interest.

abstract: |
  \textbf{Abstract} Social media's potential effects on well-being have received considerable research interest, but much of past work is hampered by an exclusive focus on demographics in the Global North and inaccurate self-reports of social media engagement. We describe associations linking 72 countries' Facebook adoption to the well-being of 946,798 individuals from 2008 to 2019. We found no evidence suggesting that the global penetration of social media is associated with widespread psychological harm: Facebook adoption predicted life satisfaction and positive experiences positively, and negative experiences negatively, both between countries and within countries but over time. The observed associations were small. Facebook adoption predicted well-being more positively for younger individuals, but country-specific results were mixed. To move beyond studying aggregates and to better understand social media's roles in people's lives, and their potential causal effects, we need more transparent collaborative research between independent scientists and the technology industry. \linebreak[4]

  \textbf{Statement of relevance} Billions of individuals across the globe use social media on a regular basis. The widespread adoption of social media---most prominently Facebook---has fundamentally shifted how humans conduct their lives. Consequently, researchers, healthcare professionals, and policy makers have repeatedly raised alarms that the spread of social media may be negatively affecting individuals', particularly young ones', mental health and well-being on a global scale. We examined the plausibility of this idea by associating 72 nations' yearly Facebook adoption rates to the well-being of 946,798 individuals from 2008 to 2019. Overall, we found that increased rates of daily and monthly active Facebook users slightly predicted greater levels of well-being, when comparing countries to other countries, or countries to themselves but over time. These findings should guide debate surrounding social media towards more empirical foundations: The importance of this topic requires that independent scientists gain access to more detailed data from the technology industry in order to better determine how, when, and why modern online platforms might be affecting their users.

keywords: well-being, social media
wordcount: "`r wordcountaddin:::word_count(filename = 'ms.Rmd')`"
bibliography: "references.bib"
floatsintext: yes
linenumbers: yes
draft: no
mask: yes
figurelist: no
tablelist: no
footnotelist: no
header-includes:
  - |
    \setlength{\parskip}{0pt}
    \setlength\parindent{24pt}
csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: apa7
classoption: doc
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
always_allow_html: true
---

@@

# Plotting options
theme_set(
  theme_linedraw(
    base_size = 10
  ) +
    theme(
      panel.grid = element_blank(),
      strip.text = element_text(margin = margin(4, 4, 4, 4, "pt")),
      axis.text = element_text(size = rel(0.7)),
      axis.ticks = element_line(size = rel(0.4))
    )
)

# Function to replace extreme percentages
percent2 <- function(x, accuracy = .1) {
  x <- percent(x, accuracy = accuracy)
  x <- if_else(x == "100.0%", ">99.9%", x)
  x <- if_else(x == "0.0%", "<0.1%", x)
  x
}
```

@@

```{r coefficients-calculate}
path <- "models/coefficients.rds"
if (!file.exists(path)) {
  # Calculate contrasts for each outcome
  h_dau <- c(
    "Average" = "dau_cw = 0",
    "13-34_Female" = 
      "dau_cw + dau_cw:age1 * -1 + dau_cw:sex1 * -1 + dau_cw:age1:sex1 * 1 = 0",
    "13-34_Male" = 
      "dau_cw + dau_cw:age1 * -1 + dau_cw:sex1 * 1 + dau_cw:age1:sex1 * -1 = 0",
    "35+_Female" = 
      "dau_cw + dau_cw:age1 * 1 + dau_cw:sex1 * -1 + dau_cw:age1:sex1 * -1 = 0",
    "35+_Male" = 
      "dau_cw + dau_cw:age1 * 1 + dau_cw:sex1 * 1 + dau_cw:age1:sex1 * 1 = 0",
    "13-34" = "dau_cw + dau_cw:age1 * -1 = 0",
    "35+" = "dau_cw + dau_cw:age1 * 1 = 0",
    "Female" = "dau_cw + dau_cw:sex1 * -1 = 0",
    "Male" = "dau_cw + dau_cw:sex1 * 1 = 0",
    "Between-country" = "dau_cb = 0",
    "X by age" = "dau_cw:age1 = 0",
    "X by sex" = "dau_cw:sex1 = 0",
    "X by age by sex" = "dau_cw:age1:sex1 = 0"
  )
  h_mau <- c(
    "Average" = "mau_cw = 0",
    "13-34_Female" = 
      "mau_cw + mau_cw:age1 * -1 + mau_cw:sex1 * -1 + mau_cw:age1:sex1 * 1 = 0",
    "13-34_Male" = 
      "mau_cw + mau_cw:age1 * -1 + mau_cw:sex1 * 1 + mau_cw:age1:sex1 * -1 = 0",
    "35+_Female" = 
      "mau_cw + mau_cw:age1 * 1 + mau_cw:sex1 * -1 + mau_cw:age1:sex1 * -1 = 0",
    "35+_Male" = 
      "mau_cw + mau_cw:age1 * 1 + mau_cw:sex1 * 1 + mau_cw:age1:sex1 * 1 = 0",
    "13-34" = "mau_cw + mau_cw:age1 * -1 = 0",
    "35+" = "mau_cw + mau_cw:age1 * 1 = 0",
    "Female" = "mau_cw + mau_cw:sex1 * -1 = 0",
    "Male" = "mau_cw + mau_cw:sex1 * 1 = 0",
    "Between-country" = "mau_cb = 0",
    "X by age" = "mau_cw:age1 = 0",
    "X by sex" = "mau_cw:sex1 = 0",
    "X by age by sex" = "mau_cw:age1:sex1 = 0"
  )

  # Summarise hypotheses with pd
  hsum <- function(fit, hypothesis) {
    h1 <- hypothesis(fit, hypothesis)
    out1 <- describe_posterior(
      h1$samples, 
      centrality = "mean", 
      test = "pd"
    ) %>% 
      mutate(
        Hypothesis = h1$hypothesis$Hypothesis, 
        country = "World",
        .before = 1
      ) %>% 
      select(-CI, -Parameter) %>% 
      tibble()
    h2 <- hypothesis(fit, hypothesis, scope = "coef", group = "country")
    out2 <- describe_posterior(
      h2$samples, 
      centrality = "mean", 
      test = "pd"
    ) %>% 
      mutate(
        Hypothesis = h2$hypothesis$Hypothesis, 
        country = h2$hypothesis$Group, 
        .before = 1
      ) %>% 
      select(-CI, -Parameter) %>% 
      tibble
    bind_rows(out1, out2)
  }

  coefs <- fits %>% 
    select(-data, -model) %>% 
    mutate(h = if_else(predictor == "dau", list(h_dau), list(h_mau))) %>% 
    mutate(
      coefs = map2(
        fit, h,
        ~hsum(.x, .y)
      )
    ) %>% 
    select(-fit, -h) %>% 
    unnest(coefs) %>% 
    # Between country contrast not country-specific!
    filter(!(Hypothesis == "Between-country" & country != "World"))

@@

To better understand the plausible range of associations, we linked data tracking Facebook's global adoption with three indicators of well-being. We examined 72 countries' per capita active Facebook users in two age brackets (13-34 and 35+ years) as predictors of life satisfaction (LS), negative (NE), and positive psychological experiences (PE) at the level of years spanning 2008 to 2019. The well-being data represented 946,798 individuals' responses from the nationally representative Gallup World Poll Survey [@gallupHowDoesGallup2014]. We joined these unique datasets to conduct a descriptive study to answer three basic yet important questions: First, to what extent is Facebook adoption associated with well-being? Second, do these associations differ by age or sex. And finally, how might these associations have differed between countries? In addition, we were interested in whether the intensity of use might make a difference, and therefore conducted our analyses separately for daily active users and monthly active users.

# Methods

@@

Both datasets (Facebook and Gallup) are proprietary and we therefore could not share them with this manuscript. Our analytic code, along with synthetic datasets and an online analysis supplement, is available at <https://doi.org/10.5281/zenodo.7086277>. This study was not preregistered. Researchers can contact Facebook (ccobb@fb.com) to reproduce our analyses with the actual Facebook adoption dataset. The Gallup data are available online to subscribing institutions.

@@

Before analysing the data, we converted Facebook DAU and MAU to proportions of each country's and age group's yearly population sizes using population data from the United Nations Department of Economic and Social affairs (<https://population.un.org/wpp/Download/Standard/Population/>). No observations were removed for analyses.

We examined the association between Facebook adoption and well-being through bayesian hierarchical regression models, estimated separately for DAU and MAU, and each of the three well-being outcomes. We regressed the outcome $y$ on time (decades, centered on 2014), within-country centered DAU or MAU (in separate models), age, sex, all the interactions except time by DAU/MAU, and the between-country centered DAU/MAU. We allowed all coefficients to vary randomly across the 72 countries, except the between-country predictor. Because we aggregated the outcomes, we included to standard errors of the outcome in the model to account for the varying group sizes and resulting uncertainties in the modelled data. Formally, we specified the model as

$$
\begin{aligned} 
y_i &\sim \text{Normal}(\mu_i, \sqrt{\sigma^{2} + \text{se}_i^2}), \\
\mu_i &= \alpha_{0} + \beta_{0\text{country}[i]} + \\
&\quad\ (\alpha_{1} + \beta_{1\text{country}[i]})\text{Sex}_i + \\
&\quad\ (\alpha_{2} + \beta_{2\text{country}[i]})\text{Age}_i + \\
&\quad\ (\alpha_{3} + \beta_{3\text{country}[i]})\text{Time}_i + \\
&\quad\ (\alpha_{4} + \beta_{4\text{country}[i]})\text{DAU}^{\text{CW}}_i + \\
&\quad\ (\alpha_{5} + \beta_{5\text{country}[i]})\text{Sex}_i \times \text{Time}_i + \\
&\quad\ (\alpha_{6} + \beta_{6\text{country}[i]})\text{Age}_i \times \text{Time}_i + \\
&\quad\ (\alpha_{7} + \beta_{7\text{country}[i]})\text{Sex}_i \times \text{Age}_i + \\
&\quad\ (\alpha_{8} + \beta_{8\text{country}[i]})\text{Sex}_i \times \text{Age}_i \times \text{Time}_i + \\
&\quad\ (\alpha_{9} + \beta_{9\text{country}[i]})\text{Sex}_i \times \text{DAU}^{\text{CW}}_i + \\
&\quad\ (\alpha_{10} + \beta_{10\text{country}[i]})\text{Age}_i \times \text{DAU}^{\text{CW}}_i + \\
&\quad\ (\alpha_{11} + \beta_{11\text{country}[i]})\text{Sex}_i \times \text{Age}_i \times \text{DAU}^{\text{CW}}_i + \\
&\quad\ \alpha_{12} \text{DAU}^{\text{CB}}_i,
\\
\pmb{\beta} &\sim \text{MVN}(\pmb{0}, \Sigma).
\end{aligned}
$$

Where $y_i$ is the outcome (e.g. life satisfaction) on row $i$, $\text{se}_i$ its standard error, $\pmb{\alpha}$ are the population-level coefficients, $\beta_{0, \dots, 11\text{country}[i]}$ are the country-specific coefficients for the country indicated on row $i$, $\text{DAU}^{\text{CW}}_i$ the within-country centered year-aggregated daily (or monthly) active Facebook users, and $\text{DAU}^{\text{CB}}_i$ the between-country centered DAU (or MAU).

We conducted all data analyses with the R language for statistical computing [@rcoreteamLanguageEnvironmentStatistical2021] and estimated the models using Stan’s Hamiltonian Monte Carlo sampling via the brms R package [@burknerBrmsPackageBayesian2017; @standevelopmentteamStanModelingLanguage2021]. We used default noninformative priors, 4 HMC chains with 4,000 iterations and first 2,000 as warmup for 8,000 total iterations; we report all parameters with their posterior means and 95% credible intervals (posterior 2.5 and 97.5 percentiles; CI), and other posterior probabilities as indicated in text.

# Results

```{r fig-data}
#| include: true
#| fig.height: 4.4
#| fig.width: 3.5
#| fig.cap: "A. 72 countries' daily (top) and monthly (bottom) active Facebook users in two age brackets from 2008 to 2019. (Percentages may exceed 100\\% due to inaccurate estimation of either DAU or population size.) B. The same 72 countries' mean responses to three well-being scales in the Gallup World Poll from 2008 to 2019, separated by age category and sex. LS: Life satisfaction, NE: Negative experiences: PE: Positive experiences."
